The variability of Betelgeuse explained by surface convection

### 2024-01-30 main.tex
Bib added

At the end of 2019, Betelgeuse suffered a sudden and dramatic loss of brightness \cite{}, particularly in the blue colors. 
\cite{} has demonstrated that such event was due to the emission by the star of a cloud of dust near to the line of sight of 
Earth. In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the 
brightness distribution over the stellar disk due to the presence of large quantities of dust. While this may be seen as 
a singular event, it brings up the question whether all changes in brightness cannot be due to similar changes in the brightness 
distribution over the disk.

@@

Linear polarization in the atomic lines of the spectra of Betelgeuse, discovered by \cite{}, has been interpreted as the joint action of 
two mechanisms. First the depolarization of the continuum by atoms which absorb linearly polarized light and re-emit unpolarized light, the 
continuum photons being polarized by Rayleigh scattering. This depolarization produces signals with an azimuth symmetry over the disk that 
would  cancel out the net linear polarization on a homogeneous disk. Thus it must be combined with an inhomogenous disk to produce the net linear polarization 
signal observed. This interpretation suggested the possibility of mapping those brightness inhomogeneities. This has been done by \cite{} 
even producing 3-dimensional images of the atmosphere of Betelgeuse \cite{} when taking advantage of the different height of formation of 
different lines in the spectrum of Betelgeuse. The images produced compare very well with cotemporaneous images made with interferometric 
techniques and show clear convective patterns. Linear polarization in the atomic lines is therefore a proxy of convection, unrelated to radial 
pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the aforementioned periods 
in linear polarization, we must conclude that these periods are related to the convective activity which originates those linear polarization 
and unrelated to any radial pulsation phenomena.

@@

Betelgeuse has been observed for the last 13 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These instruments measure the polarization 
over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
this sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in the individual atomic lines of the 
spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only excepcionnaly do they 
reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
see the linear polarization signal above noise \cite{Auriere}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
occassions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
is done through a technique called Least-Squares Deconvolution (LSD) \cite{Donati} and assumes that the linear polarization signal is similar 
in all spectral lines up to scale factors in the sampling and amplitude. This technique has been successfully used in the past to measure 
magnetic field distributions over stellar surfaces and it is now used to produce images of the brightness distributions in the photosphere 
of Betelgeuse and other red superg

@@

\end{document}

>>>>

At the end of 2019, Betelgeuse suffered a sudden and dramatic loss of brightness \cite{}, particularly in the blue colors. 
\cite{Miguel} has demonstrated that such event was due to the emission by the star of a cloud of dust near to the line of sight of 
Earth. In this event it was clear that a change in brightness was not due to any kind of pulsation but to a change on the 
brightness distribution over the stellar disk due to the presence of large quantities of dust. While this may be seen as 
a singular event, it brings up the question whether all changes in brightness cannot be due to similar changes in the brightness 
distribution over the disk.

@@

Linear polarization in the atomic lines of the spectra of Betelgeuse, discovered by \cite{auriere_discovery_2016}, has been interpreted as the joint action of 
two mechanisms. First the depolarization of the continuum by atoms which absorb linearly polarized light and re-emit unpolarized light, the 
continuum photons being polarized by Rayleigh scattering. This depolarization produces signals with an azimuth symmetry over the disk that 
would  cancel out the net linear polarization on a homogeneous disk. Thus it must be combined with an inhomogenous disk to produce the net linear polarization 
signal observed. This interpretation suggested the possibility of mapping those brightness inhomogeneities. This has been done by \cite{lopez_ariste_convective_2018} 
even producing 3-dimensional images of the atmosphere of Betelgeuse \citep{lopez_ariste_three-dimensional_2022} when taking advantage of the different height of formation of 
different lines in the spectrum of Betelgeuse. The images produced compare very well with cotemporaneous images made with interferometric 
techniques and show clear convective patterns. Linear polarization in the atomic lines is therefore a proxy of convection, unrelated to radial 
pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the aforementioned periods 
in linear polarization, we must conclude that these periods are related to the convective activity which originates those linear polarization 
and unrelated to any radial pulsation phenomena.

@@

Betelgeuse has been observed for the last 13 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These instruments measure the polarization 
over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
this sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in the individual atomic lines of the 
spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only excepcionnaly do they 
reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
see the linear polarization signal above noise \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
occassions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
is done through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997} and assumes that the 
linear polarization signal is similar 
in all spectral lines up to scale factors in the sampling and amplitude. This technique has been successfully used in the past to measure 
magnetic field distributions over stellar surfaces and it is now used to produce images of the brightness distributions in the photosphere 
of Betelgeuse and other red supergiants as cited in the Introduction.

@@

\begin{acknowledgements}
    This work was supported by the "Programme National de Physique Stellaire" (PNPS) of CNRS/INSU co-funded by CEA and CNES.
    We acknowledge support from the French National Research Agency (ANR)
    funded project PEPPER (ANR-20-CE31-0002)
    \end{acknowledgements}

    \bibliographystyle{aa}
    %\bibliographystyle{aa}

    \bibliography{art75}
\end{document}
